---
title: "ggplot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### This notebook aims to explore working with ggplot2 package, ggplot enables fast and efficient plotting. 

```{r warning=FALSE}
# Loading the packages
library(tidyverse)
library(dplyr)
library(tibble)
library(ggplot2)
```


```{r}
# Importing covid testing data using tibble read_csv function
covid_data <- read_csv('StatewiseTestingDetails.csv')
```

```{r}
covid_data
```

```{r}
## this drops the NA values from all rows
covid_data <- na.omit(covid_data)
```


### We will use ggplot2 to view the trend in covid positive cases in delhi state
```{r}
state_data = filter(covid_data, State=='Delhi')
state_data = arrange(state_data,Date)
state_data
```

### Now that we have the data for Delhi arranged in ascending order of date, we will plot the number of cases 
```{r}
ggplot(state_data, aes(x=Date, y=Positive)) # + geom_point()
```

```{r}
ggplot(state_data) + geom_text(mapping = aes(x=Date, y=Positive, label = Positive))
```

### Saving a plot
```{r}
ggsave('Delhi.png', width = 8, height = 6)
```


### Creating double plots
#### For creating double plot we need to load an additional library _gridExtra_
#### Let's create two plots for total samples and total positive cases for a particular state

```{r}
plot1 <- ggplot(state_data) + geom_point(mapping = aes(x=Date, y=TotalSamples))
plot2 <- ggplot(state_data) + geom_point(mapping = aes(x=Date, y=Positive))
library(gridExtra)
grid.arrange(plot1, plot2, ncol = 2)
```

## Lets try more things with another dataset!!
```{r}
data(mpg)
?mpg
```

### Let's plot!!
```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = class)) + geom_point()
```

### Making the plots look more nicer!
```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = class)) + geom_point() + 
  labs(title = "Fuel Efficiency by Engine Power",
       subtitle = "Fuel economy data from 1999 and 2008 for 38 popular models of cars",
       x = "Engine power (litres displacement)",
       y = "Fuel Efficiency (miles per gallon)",
       color = "Car Type") + theme_classic()
```

### Faceting 
```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  facet_grid(year ~ cyl)
```

## More on ggplot here: 
https://uc-r.github.io/ggplot_intro 
http://r-statistics.co/Complete-Ggplot2-Tutorial-Part1-With-R-Code.html 






