---
title: "ggplot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### This notebook aims to explore working with ggplot2 package, ggplot enables fast and efficient plotting. 

```{r}
# Loading the packages
library(tidyverse)
library(dplyr)
library(tibble)
library(ggplot2)
```


```{r}
# Importing covid testing data using tibble read_csv function
covid_data <- read_csv('StatewiseTestingDetails.csv')
```
```{r}
covid_data
```
### We will use ggplot2 to view the trend in covid positive cases in delhi state
```{r}
state_data = filter(covid_data, State=='Uttarakhand')
state_data = arrange(state_data,Date)
state_data
```
### Now that we have the data for Delhi arranged in ascending order of date, we will plot the number of cases 
```{r}
ggplot(state_data) + geom_point(mapping = aes(x=Date, y=Positive))
```


```{r}
ggplot(state_data) + geom_text(mapping = aes(x=Date, y=Positive, label = Positive))
```
### Saving a plot
```{r}
ggsave('UK.png', width = 8, height = 6)
```

### Creating double plots
#### For creating double plot we need to load an additional library _gridExtra_
#### Let's create two plots for total samples and total positive cases for a particular state

```{r}
plot1 <- ggplot(state_data) + geom_point(mapping = aes(x=Date, y=TotalSamples))
plot2 <- ggplot(state_data) + geom_point(mapping = aes(x=Date, y=Positive))
library(gridExtra)
grid.arrange(plot1, plot2, ncol = 2)
```



