---
title: "Infeerntial Statistics: Probability & Distributions - 2"
output: html_notebook
---

In this notebook, we'll be looking at a few more distributions. 

##### Normal Distribution
The most common one is Normal Distribution or the "Bell curve" or a "Gaussian distribution". It is described using two parameters, the mean of the distribution μ and the standard deviation of the distribution σ. The notation that we sometimes use to say that a variable X is normally distributed is as follows:
`X ~ Normal(μ, σ)`

The functions for generating normal distribution are similar to the Binomial Distribution:

- `dnorm()` - For probability density
- `pnorm()` - For cumulative probability
- `qnorm()` - For quantile of
- `rnorm()` - For random number generation

Let's generate a normal distribution having 100 data points selected randomly. Try varying this number from 10 to 1000 and check the distribution plot.

```{r}
x <- rnorm(100)
hist(x, col = 'steelblue')
abline(v = mean(x), lty = 2, lwd = 2)
```

Now let's take a look at some distributions.

##### Generating Different types of distributions

**Normal Distribution**

One of the most common distributions in statistics is the ‘normal distribution’ aka Gaussian Distribution, well known as the ‘bell curve’ due to its characteristic bell shape.

Let's generate a Normal distribution: 

```{r}
# Here we are generating a normal distribution having 100 data points selected randomly. Try varying this number from 10 to 1000 and check the distribution plot.
x <- rnorm(100)
hist(x, col = 'steelblue')
abline(v = mean(x), lty = 2, lwd = 2)
```

Poisson Distribution
```{r}
y <- rpois(100, lambda = 1.2)
hist(y, col = 'steelblue')
abline(v = mean(y), lty = 2, lwd = 2)
```
Gamma Distribution
```{r}
x_dgamma <- seq(0, 2, by = 0.04)
y_dgamma <- dgamma(x_dgamma, shape = 6)
plot(y_dgamma)
```
mean as model
```{r}
z <- rnorm(50, mean = 5, sd = 2)
mean(z)
sd(z)
quantile(z)
quantile(z, 0.16)
mean(x) - sd(x)
mean(x) + sd(x)
```
Checking Normality of a distribution

```{r}
CompanyABCProfit<-read.csv("CompanyABCProfit.csv")
attach(CompanyABCProfit)
install.packages("dplyr")
install.packages("ggpubr")
library("ggpubr")
ggdensity(Profit,main="Profit per year", xlab="YEARLY PROFIT IN INR" )
ggqqplot(Profit)
ggplot(CompanyABCProfit) +     geom_point(mapping = aes(x = Year, y = Profit))
shapiro.test(Profit)
```

Generating a distribution and checking normality of the distribution
```{r}
N<- rnorm(100, mean=2, sd= 1.3 )
P <- rpois(100, lambda = 1.2)
#generate distribution z= p*N + (1-p)P
Y = 0.1*N
z= Y+P
#test Normality
shapiro.test(z) 
plot(z)
ggdensity(z)
ggqqplot(z)
```

```{r}
u=runif(40, min = -1, max = 1)
N<- rnorm(100, mean=2, sd= 1.3 )
#generate distribution z= p*N + (1-p)u where 0<p<1
s = 0.1*N
j=0.9*u
q= s+j
#test Normality
shapiro.test(q) 
plot(q)
ggdensity(q)
ggqqplot(q)
boxplot(q)
```
Emotional Valence Data
```{r}
# Load tidyverse and Warriner et al. (2013) data:

library(tidyverse)

war <- read_csv('warriner_2013_emotional_valence.csv')

# Check:

war

# Check valence measure range:

range(war$Val)

# Check the least and most positive wors:

filter(war, Val == min(Val) | Val == max(Val))

# Same thing, but more compact:

filter(war, Val %in% range(Val))

# Check tibble in ascending order:

arrange(war, Val)

# And descending order:

arrange(war, desc(Val))

#Mean and SD:

mean(war$Val)
sd(war$Val)

# 68% rule:

mean(war$Val) + sd(war$Val)
mean(war$Val) - sd(war$Val)

# Confirm:

quantile(war$Val, c(0.16, 0.84))

# Median:

median(war$Val)

# Which is the same as the 50th percentile:

quantile(war$Val, 0.5)
```

```{r}
ggdensity(war$Val)
boxplot(war$Val)
```
1. Generate an alpha distribution.
2. Generate a normal distribution with 150 data points and mean=2, standard deviation= 1.5 and a uniform distribution of maximum value 1, minimum value -1. Generate a distribution combining both using the equation, X= p*N + (p-1)u, where N= normal distribution, u= uniform distribution. X is generated new distribution. Vary the value of p= 1,2,3.. and check normality of distribution X. 
3. Load the emotional valence dataset from warrinar et al. 2013 check mean valence, plot boxplot with indication of first and third quantile value int it. Check normality of the data. 


